KidsAIR initial SUMs data work

### 2020-01-10 sums_data_cleaning.Rmd
SUMs cleaning updates. Add extensive notes to PM and SUMs code.

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 9 Jan 2020"
output: pdf_document
---

@@

# Add linked IDs to WMT SUMs data logs
```{r}
wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = `REPLACE(REPLACE(Notes`) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Add linked IDs to NN SUMs data logs
```{r}
nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Add linked IDs to AK SUMs data logs
```{r}
ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Initial load, format, and save KidsAIR WMT SUMs data
## Files here are in several formats: csv (3 different formats) and txt (5 formats)
## I had to do some manual cleaning of the files to get them named
## consistently and in reasonably similar formats to be able to load
## into R in batches, as seen below. The files I upload in the code below
## will also be uploaded to Box in a folder with my name.
## Box folder directory: 
```{r}
##### Read in sums csv files (from WMT) #####

# List files
list_csv_files <- list.files("Input/WMT/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/WMT/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("line", "date", "time", "temp_c", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         file_type = "logtag") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")

# Set working directory and load files in list; extract file name and add as column
## run next 7 lines together
setwd("Input/WMT/sums_csv_old/format_1")
initial_data = tibble(file_list = list_csv_files_2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 22,
                            col_names = c("datetime_sums", "temp_c_sums"))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv_2 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA",
         datetime_sums = ymd_hms(datetime_sums)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(date_sums = ymd(date_sums),
         file_type = "temp_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/WMT/sums_csv_old/format_2")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 2,
                            col_names = c("line", "date", "time", "temp_f", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_double(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_sums_csv_3 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = mdy_hms(datetime),
         date_sums = mdy(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8,
         file_type = "temp_f") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

@@

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")

@@

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")

@@

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx2 <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")

@@

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx3 <- initial_data %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")

@@

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx4 <- initial_data %>% 
  mutate(temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = temp_c,
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")

@@

# Format combined file from above; combine with other PM data below
wmt_sums_xlsx5 <- initial_data %>% 
  mutate(temp_f = as.numeric(temp_f),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8,
         home = as.factor(home),
         file_type = "xlsx_f") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# Bind all WMT SUMs data together
wmt_sums_raw <- rbind(wmt_sums_csv, wmt_sums_csv_2, wmt_sums_csv_3,
                      wmt_sums_xlsx, wmt_sums_xlsx2, wmt_sums_xlsx3,
                      wmt_sums_xlsx4, wmt_sums_xlsx5) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
                                                  "temp_f", "xlsx_c",
                                                  "xlsx_f")))

@@

write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
```

# Initial load, format, and save KidsAIR NN SUMs data
```{r}
##### Read in sums csv files (from NN) #####

# List files
list_csv_files <- list.files("Input/NN/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv,
                            col_names = c("line", "date", "time", "temp_c"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_sums_csv <- initial_data %>% 
  mutate(area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c_sums = temp_c,
         file_type = "logtag") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")

@@

# Format combined file from above; combine with other PM data below
nn_sums_csv_2 <- initial_data %>% 
  mutate(area = "NN",
         datetime_sums = mdy_hm(datetime)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>%
  mutate(date_sums = ymd(date_sums),
         temp_c_sums = as.numeric(temp_c),
         file_type = "temp_c") %>%
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# List files
list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_old/3col-f")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("date", "time", "temp_f"),
                            col_types = cols(date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_double()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_sums_csv_3 <- initial_data %>% 
  mutate(area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = mdy_hms(datetime),
         date_sums = mdy(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8,
         file_type = "temp_f") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

>>>>

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 10 Jan 2020"
output: pdf_document
---

@@

# Add linked IDs to WMT SUMs data logs
## Use these throughout the data cleaning process to check start times
## Files and notes can be found on Box: 
### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
```{r}
# Input file from Box: ibutton.xlsx
wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

# Input file from Box: ibuttonupd.xlsx
wmt_sums_data_log_updated <- read_csv("Input/WMT/wmt_sums_data_log_updated.csv") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = `REPLACE(REPLACE(Notes`) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Add linked IDs to NN SUMs data logs
## Use these throughout the data cleaning process to check start times
## Files and notes can be found on Box: 
### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
```{r}
# Input file from Box: ibutton.xlsx
nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

# Input file from Box: ibuttonupd.xlsx
nn_sums_data_log_updated <- read_xlsx("Input/NN/nn_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Add linked IDs to AK SUMs data logs
## Use these throughout the data cleaning process to check start times
## Files and notes can be found on Box: 
### NoonanGroupData > KidsAIR > Ethan - data cleaning and analysis
```{r}
# Input file from Box: ibutton.xlsx
ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log)

@@

# Input file from Box: ibutton.xlsx
ak_sums_data_log_updated <- read_xlsx("Input/AK/ak_sums_data_log_updated.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         date = ChangeDate,
         time = ChangeTime,
         alarms = Alarms,
         file_name = ibutFile,
         notes = IgnoreReason) %>% 
  select(home_winter_id, date, time, alarms, file_name, notes)

@@

# Initial load, format, and save KidsAIR WMT SUMs data
## Files were downloaded from Box: KidsAIR > 2019-Nov-Data > WMT > uploads > IButton
## Files were not named consistently, so I renamed them by their home ID number.
## Files also had several different types and formats.
## Manual cleaning of some files had to be done to get them in 
## consistent formats and file types to work with in R.
## Unfortunately, this process had to be manual due to the large number of files
## and inconsistent QC process across the various files.
## As a result, this process will be difficult to replicate if needed;
## The code below will highlight how the files were formatted and grouped together.
## Once files were in consistent formats, I load them to R in batches, as seen below.
## The following code cleans and combines all SUMs temperature data
## Notes and cleaned data files are uploaded to Box:
### KidsAIR > Ethan - data cleaning and analysis
```{r}
##### Read in sums csv files (from WMT) #####

# List files
## These files were directly exported by me from the Logtag software.
## I am confident in their formatting and content since I pulled the data
## from the original file type of the field instrument.
## The files have 5 columns: line, date, time, temp (C), notes
list_csv_files <- list.files("Input/WMT/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 13 lines together
setwd("Input/WMT/sums_csv_new") # set new working directory
initial_data = tibble(file_list = list_csv_files) %>% # use list of files from above
  extract(file_list, "home", remove = FALSE) %>% # extract file name, name "home"
  # apply read_csv function; name and format columns
  mutate(save_data = lapply(file_list, read_csv, skip = 1, 
                            col_names = c("line", "date", "time", "temp_c", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above
wmt_sums_csv <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  # separate temp to remove instances of ">" in the temp data
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         file_type = "logtag") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Only 1 file had this format: 
## a 21 row header, then columns of datetime and temp (C)
list_csv_files_2 <- list.files("Input/WMT/sums_csv_old/format_1")

# Set working directory and load files in list; extract file name and add as column
## run next 9 lines together
setwd("Input/WMT/sums_csv_old/format_1")
initial_data = tibble(file_list = list_csv_files_2) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 22,
                            col_names = c("datetime_sums", "temp_c"),
                            col_types = cols(datetime_sums = col_character(),
                                             temp_c = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above
wmt_sums_csv_2 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA",
         datetime_sums = ymd_hms(datetime_sums)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
  mutate(date_sums = ymd(date_sums),
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         file_type = "temp_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Numerous files had this format: 
## 5 columns: line, date, time, temp (F), notes
list_csv_files_3 <- list.files("Input/WMT/sums_csv_old/format_2")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/WMT/sums_csv_old/format_2")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 2,
                            col_names = c("line", "date", "time", "temp_f", "trash"),
                            col_types = cols(line = col_double(),
                                             date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_character(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above
wmt_sums_csv_3 <- initial_data %>% 
  mutate(area = "WMT",
         home2 = "WMA") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_f != "Paused") %>% 
  separate(temp_f, c("trash2", "temp_f"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = mdy_hms(datetime),
         date_sums = mdy(date),
         time_sums = time,
         temp_f = if_else(is.na(temp_f), trash2, temp_f),
         temp_f = as.numeric(temp_f),
         temp_c_sums = (temp_f-32)/1.8,
         file_type = "temp_f") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

@@

# List files
## Numerous files had this format with a 16 row header
## 2 columns: datetime, temp (C)
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip16")

@@

# Format combined file from above
wmt_sums_xlsx <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## 6 files had this format with a 21 row header
## 2 columns: datetime, temp (C)
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/header_skip21")

@@

# Format combined file from above
wmt_sums_xlsx2 <- initial_data %>% 
  mutate(datetime = ymd_hms(`Log data: date/time`),
         temp_c = as.numeric(`Temperature °C`),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Numerous files had this format with no header
## 2 columns: datetime, temp (C)
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_2col_c")

@@

# Format combined file from above
wmt_sums_xlsx3 <- initial_data %>% 
  mutate(datetime = ymd_hms(datetime),
         temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Numerous files had this format with no header
## 4 columns: date, time, am/pm, temp (C)
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_c")

@@

# Format combined file from above
wmt_sums_xlsx4 <- initial_data %>% 
  mutate(temp_c = as.numeric(temp_c),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>%
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         home = as.factor(home),
         file_type = "xlsx_c") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Numerous files had this format with no header
## 4 columns: date, time, am/pm, temp (F)
list_xlsx_files <- list.files("Input/WMT/sums_xlsx_files/no_header_4col_f")

@@

# Format combined file from above
wmt_sums_xlsx5 <- initial_data %>% 
  mutate(temp_f = as.numeric(temp_f),
         area = "WMT",
         home2 = "WMA") %>% 
  separate("time", c("trash", "time"), sep = " ") %>%
  unite("time", c("time", "am_pm"), sep = " ") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  unite("home", c("home2", "home"), sep = "") %>% 
  mutate(datetime_sums = ymd_hms(datetime),
         date_sums = ymd(date),
         time_sums = time,
         temp_c_sums = (temp_f-32)/1.8,
         home = as.factor(home),
         file_type = "xlsx_f") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# Bind all WMT SUMs data together
wmt_sums_raw <- rbind(wmt_sums_csv, wmt_sums_csv_2, wmt_sums_csv_3,
                      wmt_sums_xlsx, wmt_sums_xlsx2, wmt_sums_xlsx3,
                      wmt_sums_xlsx4, wmt_sums_xlsx5) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% # remove duplicate lines that have same info from different files
  mutate(file_type = factor(file_type, levels = c("logtag", "temp_c",
                                                  "temp_f", "xlsx_c",
                                                  "xlsx_f")))

@@

# write_rds(wmt_sums_filter, "Output/wmt_sums_raw.rds")
```

# Initial load, format, and save KidsAIR NN SUMs data
## This chunk performs the same process that was done previously for WMT files
```{r}
##### Read in sums csv files (from NN) #####

# List files
## These files were directly exported by me from the Logtag software.
## I am confident in their formatting and content since I pulled the data
## from the original file type of the field instrument.
## The files have 5 columns: line, date, time, temp (C), notes
list_csv_files <- list.files("Input/NN/sums_csv_new")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_new")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv,
                            col_names = c("line", "date", "time", "temp_c", "trash"),
                            col_types = cols(line = col_character(),
                                             date = col_date(format = ""),
                                             time = col_time(format = ""),
                                             temp_c = col_character(),
                                             trash = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above
nn_sums_csv <- initial_data %>% 
  mutate(area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_sums = datetime,
         date_sums = date,
         time_sums = time,
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         file_type = "logtag") %>% 
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
# 1 file had this format: No header, columns of datetime and temp (C)
list_csv_files_2 <- list.files("Input/NN/sums_csv_old/2col-c")

@@

# Format combined file from above; combine with other PM data below
nn_sums_csv_2 <- initial_data %>% 
  mutate(area = "NN",
         datetime_sums = mdy_hm(datetime)) %>% 
  separate("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>%
  filter(temp_c != "Paused") %>% 
  separate(temp_c, c("trash2", "temp_c"), sep = " ", remove = FALSE) %>% 
  mutate(date_sums = ymd(date_sums),
         temp_c = if_else(is.na(temp_c), trash2, temp_c),
         temp_c_sums = as.numeric(temp_c),
         file_type = "temp_c") %>%
  select(home, area, datetime_sums, date_sums, time_sums, temp_c_sums, file_type) %>% 
  arrange(home, datetime_sums) %>% 
  distinct() %>% 
  filter(!is.na(temp_c_sums))

# List files
## Numerous files had this format
## No header, 3 columns: date, time, temp (F)
list_csv_files_3 <- list.files("Input/NN/sums_csv_old/3col-f")

# Set working directory and load files in list; extract file name and add as column
## run next 12 lines together
setwd("Input/NN/sums_csv_old/3col-f")
initial_data = tibble(file_list = list_csv_files_3) %>%
  extract(file_list, "home", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv, skip = 1,
                            col_names = c("date", "time", "temp_f"),
                            col_types = cols(date = col_character(),
                                             time = col_time(format = ""),
                                             temp_f = col_character()))) %>%
  unnest(save_data) %>%
  select(-file_list)
